Dominance Regions for Imaginary $\bfg$-vectors

### 2020-06-10 dominance_regions.tex
Stability of imaginary cones under mutations

  Define $\phi_k:\ZZ^n\to\ZZ^n$ as the piecewise-linear map $\phi_k(\lambda)=\begin{cases} E_{k,+}\lambda & \text{if $\lambda_k\ge0$;}\\ E_{k,-}\lambda & \text{if $\lambda_k<0$.} \end{cases}$.
  Define the piecewise-linear automorphisms $\phi_t:\ZZ^n\to\ZZ^n$ by
  \begin{itemize}
    \item $\phi_{t_0}=\Id$;
    \item if $t,t'\in\TT_n$ are joined by an edge labelled $k$, then $\phi_{t'}=\phi_k \phi_t$.
  \end{itemize}
  For $\lambda\in\ZZ^n$, define the \emph{$B$-cone} $\cC(\lambda):=\lambda+B\cdot\RR_{\ge0}^n$ pointed at $\lambda$.
  \begin{definition}
    For $\lambda\in\ZZ^n$, define the \emph{dominance polytope} 
    \[ \cP(\lambda) = \bigcap_{t\in\TT_n} \phi_t^{-1} \cC(\phi_t \lambda). \]
  \end{definition}
  Our goal is to understand the dominance polytopes $\cP(\lambda)$ for all $\lambda$.

  \section{Rank Two}
  Fix integers $b,c>0$ and assume $bc\ge 4$.
  Let $B=\left[ \begin{array}{cc} 0 & -b \\ c & 0 \end{array} \right]$.
  Here we may take $\TT_2$ to have vertices labelled $t_\ell$ for $\ell\in\ZZ$, with $t_{2r}$ joined to $t_{2r+1}$ by an edge labelled $1$ and $t_{2r-1}$ joined to $t_{2r}$ by an edge labelled $2$.
  We aim to understand the dominance regions $\cP(\lambda) = \bigcap_{i\in\ZZ} \phi_{t_i}^{-1} \cC(\phi_{t_i} \lambda)$ for $\lambda\in\ZZ^2$.

@@

  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} spanned by the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$.
  \begin{lemma}
    The vectors $(2b, -bc+\sqrt{bc(bc-4)})$ (resp. $(2b, -bc-\sqrt{bc(bc-4)})$) and $(2c, bc+\sqrt{bc(bc-4)})$ (resp. $(2c, bc-\sqrt{bc(bc-4)})$) are perpendicular.
  \end{lemma}

>>>>

  Define $\phi_k:\ZZ^n\to\ZZ^n$ as the piecewise-linear map $\phi_k(\lambda)=\begin{cases} E_{k,+}\lambda & \text{if $\lambda_k\ge0$;}\\ E_{k,-}\lambda & \text{if $\lambda_k<0$.} \end{cases}$.
  Define the piecewise-linear automorphisms $\phi_t:\ZZ^n\to\ZZ^n$ by
  \begin{itemize}
    \item $\phi_{t_0}=\Id$;
    \item if $t,t'\in\TT_n$ are joined by an edge labelled $k$, then $\phi_{t'}=\phi_k \phi_t$.
  \end{itemize}
  For $\lambda\in\ZZ^n$, define the \emph{$B_t$-cone} $\cC_t(\lambda):=\lambda+B_t\cdot\RR_{\ge0}^n$ pointed at $\lambda$.
  \begin{definition}
    For $\lambda\in\ZZ^n$, define the \emph{dominance polytope} 
    \[ \cP(\lambda) = \bigcap_{t\in\TT_n} \phi_t^{-1} \cC_t(\phi_t \lambda). \]
  \end{definition}
  Our goal is to understand the dominance polytopes $\cP(\lambda)$ for all $\lambda$.

  \section{Rank Two}
  Fix integers $b,c>0$ and assume $bc\ge 4$.
  Let $B=\left[ \begin{array}{cc} 0 & -b \\ c & 0 \end{array} \right]$.
  Here we may take $\TT_2$ to have vertices labelled $t_\ell$ for $\ell\in\ZZ$, with $t_{2j}$ joined to $t_{2j+1}$ by an edge labelled $1$ and $t_{2j-1}$ joined to $t_{2j}$ by an edge labelled $2$.
  We aim to understand the dominance regions $\cP(\lambda) = \bigcap_{i\in\ZZ} \phi_{t_i}^{-1} \cC(\phi_{t_i} \lambda)$ for $\lambda\in\ZZ^2$.

@@

  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} spanned by the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$.
  Also, write $\cI' \subset \RR^2$ for the \emph{dual imaginary cone} spanned by the vectors $(-bc\pm\sqrt{bc(bc-4)},2c)$.
  \begin{lemma}
    The vectors $(2b, -bc+\sqrt{bc(bc-4)})$ (resp. $(2b, -bc-\sqrt{bc(bc-4)})$) and $(2c, bc+\sqrt{bc(bc-4)})$ (resp. $(2c, bc-\sqrt{bc(bc-4)})$) are perpendicular.
  \end{lemma}
  \begin{proof}
    This is immediate from the equality $(-bc+\sqrt{bc(bc-4)})(bc+\sqrt{bc(bc-4)})=-4bc$.
  \end{proof}
  \begin{lemma}
    For $j\in\ZZ$ and $\lambda\in\cI$, we have $\phi_{t_{2j}}(\lambda)\in\cI$ and $\phi_{t_{2j+1}}(\lambda)\in\cI'$.
  \end{lemma}
  \begin{proof}
    Note that $\phi_{t_{2j}}=\phi_{t_2}^j$ and $\phi_{2j+1}=\phi_{t_1}\phi_{t_2}^j$ so the result follows from the cases $j=1$ and $j=0$ respectively.
    The claim is immediate for $\phi_{t_2}$ since the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$ are eigenvectors.
    Now, we have $\phi_{t_1}(\lambda)=(-\lambda_1,c\lambda_1+\lambda_2)$ and so $\phi_{t_1}\big( (2b,-bc\pm\sqrt{bc(bc-4)}) \big)=(-2b,bc\pm\sqrt{bc(bc-4)})$, which is the $\frac{bc\pm\sqrt{bc(bc-4)}}{2c}$-multiple of the boundary vector for $\cI'$ and the result follows.
  \end{proof}
